---
title: "06_analysis_2"
author: "Group 29: Helena Møller, Emma Christensen, Frederik Holmstrup, Casper Harreby & Laura Christiansen"
format: html
---

# Descriptive analysis

## Loading Libraries

```{r}
#| eval: true
library("tidyverse") 
library("ggplot2") 
library("dplyr") 
library("readr") 
library("here")
```

## Import the dataset

```{r}
# Load the dataset   
dataset <- read_tsv("../data/03_dat_aug.tsv")
```

## Age distribution of study population

```{r}

# Calculate the mean of AGE
mean_age <- mean(data$AGE, na.rm = TRUE)

data <- data |>  filter(!is.na(AGE) & AGE >= 0) # Exclude missing or negative ages

# Visualizing age distribution of study population
p1 <- ggplot(data, aes(x = AGE)) +      
  geom_histogram(fill = "lightblue", 
                 color = "darkblue", 
                 bins = 30, 
                 alpha = 0.7) + 
  labs(title = "Age Distribution of the Study Population",             
       x = "Age (years)", 
       y = "Number of People") +      
  theme_minimal(base_size = 14) + # Clean minimal theme with larger text      
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
        axis.title.x = element_text(face = "bold", size = 14), # Bold x-axis title              
        axis.title.y = element_text(face = "bold", size = 14), # Bold y-axis title         
        panel.grid.minor = element_blank()) + # Removes minor grid lines                
  scale_x_continuous(breaks = seq(0, 100, by = 10), 
                     limits = c(0, 100)) + # Set x-axis limits and breaks      
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                     breaks = seq(0, 3000, by = 500)) + # Avoids crowded y-axis and setting breaks    
  geom_segment(x = mean_age, # Static value, not in aes()
               xend = mean_age, 
               y = 0, 
               yend = 2900, 
               color = "darkgreen", 
               linetype = "dashed", 
               size = 1.0) + # Adds mean age as a vertical dashed line      
  annotate("label", 
           x = mean_age + 12, # Move label to the right of the mean                      
           y = 2400, # Position slightly below the peak                      
           label = paste("Mean Age:", round(mean_age, 1), "years"),                 
           color = "white",                 
           fill = "darkgreen",                 
           fontface = "bold",                 
           size = 4,                      
           label.padding = unit(0.4, "lines")) # Mean age shown in a text box with padding  

# Display the plot
print(p1)

# Saving the plot as a PNG file
ggsave("../results/03_description_plot_1.png", plot = p1, width = 8, height = 4, dpi = 300, bg = "white")
```

## Age distribution of stroke subtypes (STYPE)

The stroke subtypes (STYPE) are here distributed by age and visualized with boxplots. Here's a brief description of the subtypes:

**LACS (Lacunar Stroke Syndrome)**

Blockage of small penetrating arteries in deep brain structures, such as the basal ganglia, thalamus, or internal capsule. The prognosis is genereally better than other subtypes; smaller infarct size.

**PACS (Partial Anterior Circulation Syndrome)**

Partial occlusion in the anterior cerebral circulation, typically affecting the middle cerebral artery (MCA) or anterior cerebral artery (ACA). The prognosis is moderate; larger infarct than LACS, but less severe than TACS.

**POCS (Posterior Circulation Syndrome)**

Infarction in the posterior cerebral circulation, affecting the brainstem, cerebellum, or occipital lobes. The prognosis is variable, depending on location and size of the infarction.

**TACS (Total Anterior Circulation Syndrome)**

Complete occlusion of the anterior cerebral circulation, typically affecting the middle cerebral artery (MCA) or internal carotid artery (ICA). The prognosis is poor; associated with large infarcts, severe disability, and high mortality rates.

```{r}

p2<-ggplot(dataset, aes(x = STYPE, y = AGE, fill = STYPE)) +   geom_boxplot(alpha = 0.7, outlier.shape = NA) + 
  # Boxplot without outlier points   
  labs(     title = "Stroke subtypes distributed by age",     x = "Stroke Type (STYPE)",     
   y = "Age (years)"   ) +   
  theme_minimal(base_size = 14) + 
  theme(     plot.title = element_text(hjust = 0.5, face = "bold", size = 18),   
             axis.title.x = element_text(face = "bold", size = 14),   
             axis.title.y = element_text(face = "bold", size = 14),
             legend.position = "right"   )

# Display the plot
print(p2)

# Saving the plot as a PNG file
ggsave("../results/03_description_plot_2.png", plot = p2, width = 8, height = 4, dpi = 300, bg = "white")
```

```{r}
STYPE_median <- dataset |>
  group_by(STYPE) |>
  summarize(Median_AGE = round(median(AGE, na.rm = TRUE), 0))

print(STYPE_median)
```

The median age of the different stroke subtypes are shown, where the most severe subtype TACS with high mortality rates, has the highest median age of 76 years.

# Data visualisation

## Boxplot for predicted probability of death at 6 months compared by gender

```{r}

# Boxplot of predicted probability of death at 6 months distributed by gender
p3<-ggplot(dataset, aes(x = SEX, y = EXPD6, fill = SEX)) +   
  geom_boxplot(alpha = 0.7, color = "black") +    
  scale_fill_manual(values = c("darkred", "blue")) + # Custom fill colors for gender   
  labs(
    title = "Predicted probability of death at 6 months",     
    x = "Gender",     
    y = "EXPD6",     
    fill = "Gender"
  ) + 
  theme_minimal(base_size = 14) +   
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title.x = element_text(face = "bold", size = 14), 
    axis.title.y = element_text(face = "bold", size = 14), 
    legend.position = "top"
  ) +   
  scale_y_continuous(
    limits = c(0, 1), # Adjust the range to display outliers     
    expand = expansion(mult = c(0.05, 0.1))
  )

# Display the plot
print(p3)

# Saving the plot as a PNG file
ggsave("../results/03_description_plot_3.png", plot = p1, width = 8, height = 4, dpi = 300, bg = "white")
```

## Comparison of median of EXPD6 by gender

```{r}
#Show median EXPD6 by gender  

median_table <- dataset |>
  group_by(SEX) |>
  summarize(Median_EXPD6 = median(EXPD6, na.rm = TRUE)) 

print(median_table)
```

From the medians of the EXPD6 variable, it can be seen by comparing the genders, that females have a higher median of 0.73175 than males (0.53940), thus a relatively higher predicted probability of death at 6 months from randomization.
